TBD

### 2023-03-02 outputs/paper/paper3.qmd
Exploring relationships in data

---
title: "TBD"
title-block-banner: true
abstract: "TBD"
author: "Michaela Drouillard and Christina Wei"
date-format: long
format: pdf
editor: visual
toc: true
number-sections: true
bibliography: ref.bib
---

@@

cleaned_finance_survey = read_csv(
  file = here("outputs/data/cleaned_finance_survey.csv"),
  show_col_types = FALSE
)
```

>>>>

---
title: "TBD"
title-block-banner: true
abstract: "TBD"
thanks: "Code and data supporting this analysis is available at: https://github.com/christina-wei/INF3104-3-US-Survey.git"
author: 
  - Michaela Drouillard
  - Christina Wei
date: today
date-format: long
format: pdf
editor: visual
toc: true
number-sections: true
bibliography: ref.bib
---

@@

cleaned_finance_survey = read_csv(
  file = here("outputs/data/cleaned_finance_survey.csv"),
  show_col_types = FALSE
)

combined_data = 
  merge(
    cleaned_respondent_info,
    cleaned_finance_survey,
    by = c("year", "id")
  )

rm(cleaned_respondent_info)
rm(cleaned_finance_survey)
```

## Introduction

TBD

## Data

Data was cleaned and analyzed using the open source statistically programming language R [@r], using functionalities from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `kableExtra` [@rKableExtra] and `knitr` [@rKnitr]. Details of the data extraction and cleaning processes are discussed in the subsections below.

### Surveys Related to Financial Attitudes

### Respondent Demographics

### Measurement Bias

## Findings

## Explore 2021 data

survey_data_2021 =
  combined_data |>
  filter(year == 2021)

#survey_data_2021 = merge(
#  filter(cleaned_respondent_info, year == 2021),
#  filter(cleaned_finance_survey, year == 2021),
#  by = c("year", "id")
#)
```

```{r}
#| message: false
#| echo: false
#| 
survey_data_2021$financial_satisfaction =
  factor(survey_data_2021$financial_satisfaction,
         levels = c(
           "Not Satisfied At All",
           "More or Less Satisfied",
           "Pretty Well Satisfied"
          )
        )

survey_data_2021$gender =
  factor(survey_data_2021$gender,
         levels = c(
           "Male",
           "Female"
          )
        )

survey_data_2021$degree =
  factor(survey_data_2021$degree,
         levels = c(
           "Less Than High School",
           "High School",
           "Associate/Junior College",
           "Bachelors",
           "Graduate"
          )
        )

survey_data_2021$age_band =
  factor(survey_data_2021$age_band,
         levels = c(
           "18-29",
           "30-39",
           "40-49",
           "50-64",
           "65-79",
           ">=80"
          )
        )

survey_data_2021$social_class =
  factor(survey_data_2021$social_class,
         levels = c(
           "Lower Class",
           "Working Class",
           "Lower Middle Class",
           "Middle Class",
           "Upper Middle Class",
           "Upper Class"
          )
        )
```

### Financial Satisfaction

Overall spread

```{r}
#| message: false
#| echo: false
#| warning: false

# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
survey_data_2021 |>
  ggplot(aes(x = financial_satisfaction)) +
  geom_bar(aes(y = (..count..) / sum(..count..))) +
  scale_y_continuous(labels=scales::percent)

```

Female less satisfied

```{r}
#| message: false
#| echo: false
#| warning: false

# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
survey_data_2021 |>
  filter(!is.na(gender), !is.na(financial_satisfaction)) |>
  ggplot(aes(x = financial_satisfaction, group = gender)) +
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  ylab("Percentage of response") +
  xlab("Financial Satisfaction") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none") +
  facet_grid(~gender)
```

Higher degree - more financially satisfied

```{r}
#| message: false
#| echo: false
#| warning: false

# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
survey_data_2021 |>
  filter(!is.na(degree), !is.na(financial_satisfaction)) |>
  ggplot(aes(x = financial_satisfaction, group = degree)) +
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  ylab("Percentage of response") +
  xlab("Financial Satisfaction") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none") +
  facet_grid(~degree)
```

Those born in US are less satisfied?

```{r}
#| message: false
#| echo: false
#| warning: false

# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
survey_data_2021 |>
  filter(!is.na(born), !is.na(financial_satisfaction)) |>
  ggplot(aes(x = financial_satisfaction, group = born)) +
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  ylab("Percentage of response") +
  xlab("Financial Satisfaction") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none") +
  facet_grid(~born)
```

Older cohorts are more satisfied with their financial situation

```{r}
#| message: false
#| echo: false
#| warning: false

# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
survey_data_2021 |>
  filter(!is.na(age_band), !is.na(financial_satisfaction)) |>
  ggplot(aes(x = financial_satisfaction, group = age_band)) +
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  ylab("Percentage of response") +
  xlab("Financial Satisfaction") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none") +
  facet_grid(~age_band)

```

Some differences based on working status

```{r}
#| message: false
#| echo: false
#| warning: false

# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
survey_data_2021 |>
  filter(!is.na(work_status), !is.na(financial_satisfaction)) |>
  ggplot(aes(x = financial_satisfaction, group = work_status)) +
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  ylab("Percentage of response") +
  xlab("Financial Satisfaction") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none") +
  facet_grid(~work_status)
```

Sexual orientation - not too much of a pattern?

```{r}
#| message: false
#| echo: false
#| warning: false

# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
survey_data_2021 |>
  filter(!is.na(sexual_orientation), !is.na(financial_satisfaction)) |>
  ggplot(aes(x = financial_satisfaction, group = sexual_orientation)) +
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  ylab("Percentage of response") +
  xlab("Financial Satisfaction") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none") +
  facet_grid(~sexual_orientation)
```

Social status - interestingly, upper class is less satisfied than upper middle class?

```{r}
#| message: false
#| echo: false
#| warning: false

# Reference code from https://sebastiansauer.github.io/percentage_plot_ggplot2_V2/
survey_data_2021 |>
  filter(!is.na(social_class), !is.na(financial_satisfaction)) |>
  ggplot(aes(x = financial_satisfaction, group = social_class)) +
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") + 
  scale_y_continuous(labels=scales::percent) +
  ylab("Percentage of response") +
  xlab("Financial Satisfaction") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  theme(legend.position="none") +
  facet_grid(~social_class)
```

```{r}
#| message: false
#| echo: false
#| warning: false

financial_satisfaction_across_years =
  combined_data |>
    select (year, financial_satisfaction) |>
    group_by(year) |>
    count(financial_satisfaction)

financial_satisfaction_across_years
```

## Discussion

## Conclusion

## Reference
